# Reproducible Research: Peer Assessment 1
```{r setoptions, echo=FALSE}
opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data  

Loading activity.csv into a data frame "monitor"
```{r loading}
monitor <- read.csv("activity.csv")
```

Use tapply to get the sum of steps each day.
```{r split_date}
steps_per_day <- tapply(monitor$steps, monitor$date, sum, na.rm=T)
```

Make a histogram of the total number of steps taken each day
```{r hist_steps}
hist(steps_per_day, col = "green")
rug(steps_per_day)
```

Calculate and report the mean and median total number of steps taken per day
```{r summary_steps}
mean(steps_per_day)
median(steps_per_day)
```

## What is the average daily activity pattern?

Split steps by intervals, then use sapply to get the average of steps each interval.
```{r split_interval}
steps_per_interval <- sapply(with(monitor, split(steps, interval)), mean, na.rm=TRUE)
```

Make a time series plot of the 5-min intervals and the average number of steps taken, averages across all days.
```{r plot_interval}
plot(as.table(steps_per_interval), type = "l", col="blue")
```

Which 5-min interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r which_max}
which(steps_per_interval == max(steps_per_interval))
```
So the __104^(th) 5-minute interval__, i.e. __835__, contains the maximum number of steps.
## Imputing missing values

Calculate and report the total number of missing values in the dataset.
```{r total_na}
sum(is.na(monitor$steps))
```

Devise a strategy for filling in all of the missing values in the dataset.  
_Fill by the mean for 5-min interval_
```{r fill_na}
filler <- rep(steps_per_interval, 61)
steps <- monitor$steps
na <- which(is.na(steps))
steps[na] <- filler[na]
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r new_dataset}
monitor_imputed <- cbind(steps, monitor[, c(2, 3)])
```

Make a histogram of the total number of steps taken each day.
```{r hist_steps_rm_na}
steps_date_rm_na <- sapply(with(monitor_imputed, split(steps, date)), sum)
hist(steps_date_rm_na, col = "green")
rug(steps_date_rm_na)
```

Calculate and report the mean and median total number of steps taken each day
```{r summary_steps_rm_na}
mean(steps_date_rm_na)
median(steps_date_rm_na)
```

## Are there differences in activity patterns between weekdays and weekends?
Use strptime() to convert date into POSIXlt
```{r POSIXlt}
date <- monitor_imputed$date
date <- strptime(date, "%Y-%m-%d")
```

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r plotdata}
week <- weekdays(date)
week[week %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")] <- "weekday"
week[week %in% c("Saturday", "Sunday")] <- "weekend"
week <- factor(week)
plotdata <- cbind(monitor_imputed[c(1, 3)], week)
```

Make a panel plot containing a time series plot of the 5-min intervals and the average number of steps taken, averaged across all weekday days and weekend days.

```{r baseplot}
steps_weekday <- sapply(with(plotdata[week == "weekday", ], split(steps, interval)), mean)
steps_weekend <- sapply(with(plotdata[week == "weekend", ], split(steps, interval)), mean)
par(mfrow = c(2,1))
plot(steps_weekday, type = "l", col="blue")
plot(steps_weekend, type = "l", col="red")
```
